Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-04-23 extended_catalog/paper/extended_search.tex
reorganize
updated draft

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{extension_sensitivity}
we calculate the LAT detection threshold to resolve the extension
of a source.  In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  Therefore, in our
search for extended sources we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.

\section{Validation of the Analysis Method}
\label{monte_carlo_validation}

@@

@@

@@

\bibliography{extended_catalog}

@@

>>>>

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  
\newtext{
In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{bias_wrong_spatial_model}
we asses the bias that is is introduced by using an incorrect spatial model.
In Section~\ref{extension_sensitivity}
we calculate the LAT detection threshold to resolve the extension
of a source.}
In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  
\newtext{In section XXX, we perform a dedicated monte carlo simulation
to asses possible bias due to incorrectly modeling the spatial structure
of an extended source.}
Therefore, in our
search for extended sources we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.

\section{Validation of the Analysis Method}

\subsection{\newtext{Statistical Significance Testing}}

\label{monte_carlo_validation}

@@

\newtext{
These previous simulations were preformed assuming an isotropic backgorund.
We performed a second test only

The flux of the source was scaled as
\begin{equation}
  F(\vec{x}) = F(GC) \times \left(\frac{B(\vec{x})}{B(GC)
\end{equation}

.
The benifitted from probing a large range of fluxes and spectral indices
and by testing two energy ranges.
But were unable to assess any possible systematics that may be due to
structured diffuse emission.

Assessing possible difficulties due to having
an incorrect

We performed an additional Monte Carlo test with sources

Figure~\ref{tsext_plane_plot}

We note that this study does not asses issues due to incorrectly modeling the 
galactic diffuse background which are much more difficult to assess. But
this test does show the power of our tools when the model is good (or something).

\subsection{\newtext{Assessing the Bias Due to Incorrect Spatial Models}}
\label{bias_wrong_spatial_model}

\newtext{
We performed a Monte Carlo study to asses the biases introudced by incorreclty
modeling the spatial structure of an extended source. This is important because any assumed spatial
model is just an approximation to the true photon distribution of the source.

To assess this systematic, we simulated the spatially extended ring-type SNR W44.
W44 was slelected because it is the most signfiicant extended source detected by the LAT 
that has been reported as having a non-radially symmetric photon-distribution \cite{w44}.

W44 was simulated with a power-law spectral model. The flux and spectral index of this source
were taken from our re-analysis in this paper which is described in Section~\ref{validate_known}
and summarized in Table~\label{known_extended_sources}. In particualr W44 was simulated
with a flux of $7.12\times10^{-8}$ \phflux in the energy range from 1 \gev to 100 \gev with a spectral
index of 2.66.

W44 was simulated with an elliptical ring spatial model described in
\cite{w44}.  Because that publication does not provide the numbers describing
the parameters of the ellipse, we summarize them here for completeness. The elliptical
spatial model has a major axis of 0\fdg3 and a minor axis of 0\fdg19. The
ellipse predicts emission only for the outer 0.75\% of the ellipse and
the ellipse has an angle of $147\degree$ measured east of celestial north.

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the souce was simulated on
top of the standard galactic and isotropic diffuse background model used
in 2FGL.  This simulation was in the 1 \gev to 100 \gev energy range and
the sources were simulated over the two-year time range included in the
2FGL catalog. We performed 985 identical simulations with different seeds to
see the statistical spread in the fit values.

After simulating the ring-like source, we fit the spectral and spatial parametesr of 
the ring-like extended source with successivly more complicated spatial models.
We obtained the best fit spatial parameters using \pointlike and, using 
these best fit spatial paramters, obtained the best fit spectral paramters 
using \gtlike. In total, we performed the fit using a point-like spatial model,
using a radially-symmetric uniform Gaussian and disk spatial model, 
an Elliptical disk spatial model, and finally the Monte Carlo true Elliptical
ring spatial model.

Figure~\ref{ts_comparison_w44sim} shows the the distribution of TS values for 
each of the simluations when fitting the simulated sources 
with the different spatial models. 
Figure~\ref{ts_comparison_w44sim}a shows the distribution of \ts values when 
fitting the simulated Elliptical ring extended source as as a point-like source.
the source is very significant as a point source ($\ts\sim3500$).
Figure~\ref{ts_comparison_w44sim}b shows the distirbution of \tsext values
when fitting the source as a radially-symmetric uniform disk or Gaussian source.
The source is very significiantly extended ($\tsext\sim250$). 
On average \tsext is slightly larger when
fitting the source with the Disk spatial model compared to the Gaussian spatial model.
Considering that the simulated source's true spatial model has a sharp edge, it
makes sense that the Disk would provide a better fit.

Figure~\ref{ts_comparison_w44sim}c then shows the change in \ts (or
twice the improvement in log likelihood) when fitting the simulated source as
an elliptical disk compared to fitting the source as a disk. Since the change
in \ts is $\sim30$, we see that our simulated source is significantly better
described with an ellpitical spatial model. Finally, 
Figure~\ref{ts_comparison_w44sim}d shows the change in \ts when fitting
the source as an elliptical ring compared to fitting it as an elliptical disk.
Since the change is $\sim9$, we see that the source is somewhat better described
by the true spatial model, but not at a high statistical significance.

Figure~\ref{bias_w44sim} shows the distirbution of best fit parameters
for these simulations. Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b
show the distribututions of best fit flux and spectral indices.
They show that when fitting this simulated extended source with a point-like
spatial model, there is a $\sim$10\% bias in the fit flux and
$\sim5\%$ bias in the fit index. But when fitting the sources with any of
the extended spatial models, there is no bias in the best fit spectral parameters
(within the precision of this Monte Carlo study. The reason for this is
because within the finite angular precision and limited statsitics of the LAT, 
all of the extended spatial models can be ajusted to look almost the same.
For brighter or larger sources, this could potentially be a larger effect.

Figure~\ref{bias_w44sim}c plots the best fit 68\% contaiment radius
(\rsixeight).  For the elliptical ring spatial model, as expected
there is no significant bias in the best fit size. What is surprising
is that for none of the tested spatial models was the bias greater
than $\sim10\%$. This shows that for whatever spatial model is used,
the best fit \rsixeight is a reasonable measure of the true 68\%
containment radius.

For completeness, for the elliptical disk and ring spatial models,
Figure~\ref{bias_w44sim}c plots the distribution of best fit eccentricity
of the ellipse. It is defined as $\sqrt{1-b^2/a^2}$ where $a$ is the
length of the semi-major axis and $b$ is the length of the semi-minor
axis.  Figure~\ref{bias_w44sim}c plots the distribution of best fit
orientation angle of the ellipse. For the elliptical spatial models,
there is no systematic bias in the eccentricities or orentiation angles.
For the true sptial model, there is less spread in the distributions of
the fit parameters.  Finally, for completeness Figure~\ref{bias_w44sim}d
shows that the distirubtion of best fit ring fractions is unbiased when
the source is fit with the for the elliptical ring spatial model.

@@

\newtext{In this paragraph, we describe how our test is really the same as the AIC test
because \dots}

@@

\bibliography{extended_search}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/plot_tsext_plane_color.eps}
    \else
    \plotone{mc_plots/plot_tsext_plane_bw.eps}
    \fi
    % this plot came from $tsext_plot_plane/tsext_plane_plot.py
    \caption{
    \newtext{
    ts ext in gal plane
    }
    }\label{tsext_plane_plot}
  \end{figure}

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/ts_comparison_w44sim_color.eps}
    \else
    \plotone{mc_plots/ts_comparison_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_ts_values.py
    \caption{
    \newtext{The distirbution of \ts values when fitting statistically
    independent simulations of the
    elliptical ring shaped SNR W44. (a) is the distribution of \ts values
    when fitting W44 as a point-like source and (b) is the
    distribution of \tsext when fitting the SNR with a Gaussian and uniform disk 
    spatial model. (c) is the distribution of the change in TS when
    fitting the source with an elliptical disk spatial model compared to
    fitting it with a radially symmetric spatial model and (d) 
    when fitting the source with a elliptical ring spatial model compared
    to an elliptical disk spatial model.
    }
    }\label{ts_comparison_w44sim}
\end{figure}

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{the distribution of fit parameters
    for the simulation of Monte Carlo simulations of W44.
    The plots show the distribution of (a)
    (a) best fit flux (b) spectral index (c) 68\% containment
    radius (d) eccentricity (e) elliptical orientation angle
    (f) inner ring fraction.
    }
    }\label{bias_w44sim}
\end{figure}
